---
title: "Bass Model"
output:
  pdf_document: default
  html_document: default
date: "2025-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(knitr)
library(readxl)
library(dplyr)
library(ggpubr)
```

You can also embed plots, for example:

```{r}
data <- read_excel("./data/dating_data.xlsx", sheet=2)
data

```
###Data Manipulations
####We remove unnecessary data.

```{r}
# Remove the last 4 rows
data_clean <- data[3:(nrow(data) - 4), ]


colnames(data_clean) <- c("Year", "Online Dating", "Casual Dating", "Matchmaking")

# Convert 'Online Dating' column to numeric
data_clean$`Online Dating` <- as.numeric(data_clean$`Online Dating`)

# Keep only 'Year' and 'Online Dating' columns, and remove the others
data_clean <- data_clean %>%
  select(Year, `Online Dating`)


# View the cleaned data
print(data_clean)


```

###Visualizing Data 



```{r}

# Create a bar chart for Online Dating service users over time(2018-2024)
ggplot(data_clean, aes(x = Year, y = `Online Dating`)) +
  geom_bar(stat = "identity", fill = "pink", color = "red") +  # Bar chart with pink bars and red borders
  labs(
    title = "Online Dating Service Users from 2018 to 2024",
    x = "Year",
    y = "Number of Users (in millions)"
  ) +
  theme_minimal()  
```
###Bass Model for the Online Dating data 
```{r}

years <- c(2018, 2019, 2020, 2021, 2022, 2023, 2024)
values <- c(241.89, 278.19, 313.97, 344.76, 364.05, 381.48, 395.22)

# Define the Bass model function f(t, p, q)
bass.f <- function(t, p, q) {
  ((p + q)^2 / p) * exp(-(p + q) * t) / (1 + (q / p) * exp(-(p + q) * t))^2
}

# Define the cumulative Bass model function F(t, p, q)
bass.F <- function(t, p, q) {
  (1 - exp(-(p + q) * t)) / (1 + (q / p) * exp(-(p + q) * t))
}

# Create the t variable (time)
t <- 1:length(values)

# Fit the model using nonlinear least squares (nls)
# We use the bass.f function and estimate parameters m, p, q
model <- nls(values ~ m * bass.f(t, p, q), 
             start = list(m = sum(values), p = 0.02, q = 0.4))

# View the model summary to check the fitted parameters
summary(model)

# Extract the fitted parameters (m, p, q)
fitted_params <- coef(model)
fitted_params


```
From our model summary we can conclude than our parameters are the following:

m (market potential): 5828.00

p (coefficient for innovation): 0.0361

q (coefficient for imitation): 0.191

###Predicting the diffusion of the AI Dating Assistant using the estimated parameters.
```{r}
# Parameters from the fitted model
p <- 0.0361    # Innovation coefficient
q <- 0.191     # Imitation coefficient

# Create the time steps (t)
t <- 1:length(values)


# Plot the actual sales data (sm_sales)
sm_sales <- ggplot(data.frame(Year = years, Sales = values), aes(x = Year, y = Sales)) +
  geom_line(color = 'pink') +
  geom_point(color = 'red') +
  labs(title = "Actual Sales Data", x = "Year", y = "Sales") +
  theme_minimal()

# Plot the predicted diffusion curve based on the Bass model
time_ad = ggplot(data.frame(t = c(1:14)), aes(t)) +
  stat_function(fun = bass.f, args = c(p = p, q = q)) +
  labs(title = 'Bass Model Diffusion Curve (f(t))', x = "Time", y = "New Adopters") +
  theme_minimal()

# Combine both plots side by side
ggarrange(time_ad, sm_sales, ncol = 2)

```

###Estimating the number of Adoptors 
```{r}

users<- data.frame(
  year = c(2018, 2019, 2020, 2021, 2022, 2023, 2024),
  sales = c(241.89, 278.19, 313.97, 344.76, 364.05, 381.48, 395.22)
)

p <- 0.0361    
q <- 0.191 


# Use the Bass model to predict sales for years 1 to 7 (scaling by 5828)
users$pred_sales <- bass.f(1:7, p = p, q = q) * 5828

# Create the plot
ggplot(data = users, aes(x = year, y = sales)) +
  geom_bar(stat = 'identity', fill = 'pink', alpha = 0.6) +  # Bar plot for actual sales
  geom_point(mapping = aes(x = year, y = pred_sales), color = 'red', size = 3) +  # Predicted sales as red points
  labs(title = 'Actual vs Predicted Sales', x = 'Year', y = 'Sales') +
  theme_minimal()  # Optional for a cleaner look


```
Sources

Innovation:https://time.com/7094844/rizz/

Dataset:https://www.statista.com/forecasts/891146/eservices-dating-services-online-user-by-segment-worldwide

Article about Rizz: https://www.forbes.com/sites/josipamajic/2024/09/09/rizz-app-how-the-5th-most-downloaded-dating-app-is-redefining-digital-relationships/?utm_source=chatgpt.com
